\chapter{Data}
\label{chap:data}

\section{Data Architecture}

The empirical dataset is a monthly euro-area/Germany panel built from high-frequency ECB surprise files, ECB monetary and banking series, Eurostat indicators, market data, and official monthly proxy candidates. The canonical file, \nolinkurl{data/processed/eu_de/final_monthly_model_dataset.csv}, contains 250 observations from 2005-01-31 through 2025-10-31 and 171 variables. It is rebuilt by \nolinkurl{src/data/ecb_monetary_surprises.py} and \nolinkurl{src/data/build_monthly_model_dataset.R}.

\artifacttable
  {Canonical monthly model dataset}
  {tab:canonical-monthly-dataset}
  {../data/processed/eu_de/final_monthly_model_dataset.csv}
  {Monthly panel used by the thesis-facing reduced-form local projections, timing outputs, robustness checks, and uncertainty analysis.}

\artifacttable
  {Monthly variable dictionary}
  {tab:monthly-variable-dictionary}
  {../data/processed/eu_de/monthly_variable_dictionary.csv}
  {Auditable catalog of variable names, monthly status, identification role, and non-interpolation notes.}

\nolinkurl{scripts/run_full_pipeline.py} reconstructs shocks, processed data, diagnostics, proxy validation, LP tables, figures, robustness outputs, memos, and the final audit. Source manifests, coverage tables, and diagnostics remain the factual reference layer.

\section{Variable Selection Logic}

Variables enter only if they are traceable, reproducible, compatible with monthly LP estimation or explicitly quarter-end observed, and narrower than the thesis claim. A mortgage-credit proxy is not a house-price measure; a wage tracker is not total compensation; an administered policy rate is not the high-frequency shock.

\subsection{Monetary Policy and Liquidity Variables}

\nolinkurl{target_factor_monthly_easing} is the monthly sum of ECB target-factor announcement-window surprises, signed so positive values denote easing. Timing, forward-guidance, QE, and weighted-composite variants are robustness shocks. The ECB deposit facility rate enters as month-end and monthly-average series; first differences support bridge diagnostics, while the monthly average enters LP controls and lending-spread construction.

The Wu-Xia euro-area shadow rate is retained from 2005-01 through 2022-08 \parencite{WuXia2016}. ECB asset stocks are converted from weekly observations to month-end and monthly-average series, then transformed into logs, differences, and scaled variants.

\subsection{Lending Conditions and Credit Quantities}

Mortgage and NFC lending rates come from ECB MIR sources. The main lending-condition variables subtract \nolinkurl{dfr_mavg} to form \nolinkurl{ecb_mir_mortgage_lending_spread_dfr} and \nolinkurl{ecb_mir_nfc_lending_spread_dfr}. These spreads measure borrower-facing margins relative to the policy-rate floor.

Broad credit enters through log household and NFC loan stocks. Housing finance is measured by \nolinkurl{ecb_house_purchase_growth_yoy}, the canonical monthly house-purchase credit-growth proxy, and \nolinkurl{ln_ecb_house_purchase_pure_new_loans}, a shorter-window flow series beginning in 2017-08. ECB Bank Lending Survey variables remain quarter-end observed institutional context.

\subsection{Market, Housing-Credit, and Compensation Variables}

The market-price response is \nolinkurl{ln_dax_real_de}, constructed by deflating the DAX close with German HICP and taking logs. Quarterly German house prices are preserved as \nolinkurl{ln_house_price_de_real_q_observed}, not converted into a monthly LP response.

The canonical compensation-linked response is \nolinkurl{ecb_wage_tracker_ex_oneoffs_real_yoy}, the ECB Wage Tracker excluding one-off payments minus EA20 annual inflation. Headline, unsmoothed, moving-average, and momentum variants are robustness measures. Quarterly compensation per employee remains \nolinkurl{ln_compensation_ea20_real_q_observed}. Employment expectations, inverse unemployment tightness, CES expectations, retail volume, and the German industry wage bill provide indirect labor-pressure and real-activity diagnostics.

\section{Frequency Governance}

Quarterly interpolation is rejected because it would create artificial smoothness and mechanically allocate timing across months. Quarterly house prices, compensation per employee, and BLS indicators remain observed only at quarter ends. The processed coverage file records 83 observations for the quarterly house-price and compensation series, but no usable monthly log differences.

The accepted proxies are narrower but observed monthly: house-purchase lending growth for mortgage finance and wage-tracker variables for negotiated wage pressure. BLS variables remain sparse context.

\section{Proxy Validation Framework}

Proxy selection is implemented in \nolinkurl{scripts/run_proxy_validation.py}. Outputs classify variables as canonical, robustness, indirect, diagnostic, rejected, or unavailable.

Housing proxies are scored on conceptual proximity, monthly availability, coverage, interpretability, and contamination sensitivity. \nolinkurl{ecb_house_purchase_growth_yoy} is canonical; pure-new house-purchase loans are robustness evidence; broad household loans are background credit. Compensation proxies are scored on frequency integrity, coverage, stationarity, horizon durability, sign consistency, regime stability, bootstrap robustness, clean-event sensitivity, volatility, interpretability, and revision risk. The wage tracker excluding one-offs is canonical; the German industry wage bill is a monthly robustness proxy; employment expectations and unemployment tightness are indirect diagnostics.

The appendix reports the compensation proxy ranking figure in \Cref{fig:appendix-compensation-proxy-rankings}; it supports proxy selection and is not used as an additional results claim.

\section{Transformation Pipeline}

Positive-valued stocks and indexes are logged; first differences subtract the previous monthly observation; year-on-year growth uses twelve-month log differences; and real variables deflate nominal series with the relevant HICP index. ECB wage trackers are nominal annual growth rates, so real versions subtract EA20 annual HICP inflation. The German industry wage-bill proxy is
\begin{equation}
  100\left[\log(W_t) - \log(W_{t-12})\right] - \pi^{DE}_{t,t-12},
  \label{eq:real-wage-bill}
\end{equation}
where \(W_t\) is the monthly wage-bill index and \(\pi^{DE}_{t,t-12}\) is German annual inflation.

Lending spreads are
\begin{equation}
  \mathrm{Spread}^{j}_{t}
  =
  r^{j}_{t}
  -
  \mathrm{DFR}^{mavg}_{t},
  \label{eq:lending-spread}
\end{equation}
where \(j\) denotes mortgage or NFC lending. Response variables are not standardized; only the shock response is normalized in the LP output.

\section{Final Sample Construction}

The final monthly sample runs from 2005-01-31 to 2025-10-31. Missing values are handled by response-specific estimation samples rather than imputation. The target shock is nearly complete; the Wu-Xia shadow rate ends in 2022-08; pure-new house-purchase loans begin in 2017-08; and the wage tracker begins in 2013-01.

Event filtering is layered onto the same panel. The full archive contains 244 events; within the monthly identification window, 76 are clean and 132 are contaminated or ambiguous. COVID, crisis-window, QE-launch, extreme-shock, rolling, recursive, and regime variants are robustness devices. With the data architecture fixed, the next chapter turns to response evidence.
